Variational Inference for Inverse Reinforcement Learning with Gaussian Processes

### 2019-03-11 mpaper/paper.tex
Minor readability improvements

\subsection{Linear Algebra and Analysis}

@@

\section{Solution}
% TODO: express all lambdas as exponentials (update the derivatives or sth)

\subsection{Notation}

We use $\tr(\mathbf{A})$ to denote its \emph{trace} and
$\adj(\mathbf{A})$ for its \emph{adjugate} (or \emph{classical adjoint}).

For any vector $\mathbf{x}$, we write $\mathbb{R}_d[\mathbf{x}]$ to denote a
vector space of polynomials with degree at most $d$, where variables are
elements of $\mathbf{x}$, and coefficients are in $\mathbb{R}$.

\subsection{Preliminaries}

@@

As recently suggested by Ong et al. \cite{ong2018gaussian}, we use a
decomposition $\bm\Sigma = \mathbf{B}\mathbf{B}^\intercal + \mathbf{D}^2$, where
$\mathbf{B}$ is a lower triangular $m \times p$ matrix with positive diagonal
entries, and $\mathbf{D} = \diag(d_1, \dots, d_m)$. Typically, we would set $p$
so that $p \ll m$ to get an efficient approximation, but it is also worth
pointing out that we can retain full accuracy by setting $p = m$ and $\mathbf{D}
= \mathbf{O}$. Moreover, we define a few variables that will simplify
expressions for the derivatives:
\begin{align*}
  \mathbf{U} &= (\mathbf{u} - \bm\mu)(\mathbf{u} - \bm\mu)^\intercal \\
  \mathbf{S} &= \Kru^\intercal\Kuu^{-1}, \\
  \bm\Gamma &= \Krr - \mathbf{S}\Kru, \\
  \mathbf{R} &= \mathbf{S}\frac{\partial \Kru}{\partial \lambda_i} - \frac{\partial \Krr}{\partial \lambda_i} + \left( \frac{\partial \Kru^\intercal}{\partial \lambda_i} - \mathbf{S}\frac{\partial \Kuu}{\partial \lambda_i} \right) \Kuu^{-1}\Kru, \\
  Q &= (\mathbf{u} - \bm\mu)^\intercal\bm\Sigma^{-1}(\mathbf{u} - \bm\mu).
\end{align*}

% TODO: define PDF before this point
\begin{restatable}[Derivatives of PDFs]{lemma}{derivatives} \label{lemma:derivatives}
  \leavevmode
  \begin{enumerate}
  \item $\frac{\partial q(\mathbf{u})}{\partial \bm\mu} =
    \frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} + \bm\Sigma^{-\intercal})(\mathbf{u}
    - \bm\mu)$.
  \item
    \begin{enumerate}
    \item
      $\frac{\partial q(\mathbf{u})}{\partial \bm\Sigma} =
      \frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-\intercal}\mathbf{U}\bm\Sigma^{-\intercal}
      - \bm\Sigma^{-\intercal})$.
    \item
      $\frac{\partial q(\mathbf{u})}{\partial \mathbf{B}} =
      q(\mathbf{u})(\bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1} -
      |\bm\Sigma|^{-1}\adj(\bm\Sigma))\mathbf{B}$.
    \end{enumerate}
  \item For $i = 0, \dots, d$,
    \begin{enumerate}
    \item
      \begin{align*}
        \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} &= \frac{1}{2}q(\mathbf{r} \mid \mathbf{u}) (|\bm\Gamma|^{-1} \tr(\mathbf{R} \adj(\bm\Gamma)) \\
                                                                          &- (\mathbf{r} - \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r} - \mathbf{Su})).
      \end{align*}
    \item For any covariance matrix $\mathbf{K}$,
      \[
        \frac{\partial \mathbf{K}}{\partial \lambda_i} =
        \begin{cases}
          \frac{1}{\lambda_i}\mathbf{K} & \text{if } i = 0, \\
          \mathbf{L} & \text{otherwise,}
        \end{cases}
      \]
      where
      \[
        L_{j,k} = k(\mathbf{x}_j, \mathbf{x}_k) \left( -\frac{1}{2}(x_{j,i} -
          x_{k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2 \right).
      \]
    \end{enumerate}
  \end{enumerate}
\end{restatable}

\subsection{Details of the Model} \label{sec:details}

In order to properly investigate the difference between variational inference
and ML estimation, we keep other parts of the model the same.
Namely, we set the covariance function to a version of the automatic relevance
detection kernel \cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
\begin{align*}
  k(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left( \vphantom{\frac{1}{2}} \right. &-\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i - \mathbf{x}_j) \\
                                                     &- \mathbbm{1}[i \ne j]\sigma^2\tr(\bm\Lambda) \left. \vphantom{\frac{1}{2}} \right),
\end{align*}
where $\lambda_0$ is the overall `scale' factor for how similar or distant the
states are, $\bm\Lambda = \diag(\lambda_1, \dots, \lambda_d)$ is a diagonal
matrix that determines the relevance of each feature (where $d$ denotes the
number of features), $\mathbbm{1}$ is defined as
\[ \mathbbm{1}[b] = \begin{cases}
    1 & \mbox{if $b$ is true} \\
    0 & \mbox{otherwise},
  \end{cases} \]
and $\sigma^2$ is set to $10^{-2}/2$ (as the original paper noted that
the value makes little difference to the performance of the algorithm
\cite{DBLP:conf/nips/LevinePK11}). Our vector of parameters for the
covariance function is then $\bm\lambda = (\lambda_0, \dots,
\lambda_d)^\intercal$. % TODO: write about lambda having to be variational

TODO: introductory sentence for the following assumption:
\begin{equation} \label{eq:approximation}
  \approximation = q(\mathbf{u}) \times q(\mathbf{r} \mid \mathbf{u}),
\end{equation}
where
\begin{equation}
  q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \mathbf{m}, \mathbf{S}),
\end{equation}
where $\bm\mu \in \mathbb{R}^{m}$ is the mean vector and the $m \times m$
positive semi-definite matrix $\bm\Sigma$ is the covariance matrix
\cite{DBLP:conf/nips/ChengB17,DBLP:journals/jmlr/HensmanDS17,DBLP:conf/aaai/HoangHL17}.

  \begin{figure}
    \centering
    \begin{tikzpicture}
      \node[latent] (u) {$\mathbf{u}$};
      \node[latent, right=of u] (r) {$\mathbf{r}$};
      \node[obs, right=of r] (D) {$\mathcal{D}$};
      \node[const, below=of r] (lambda) {$\bm\lambda$};
      \node[const, below=of u] (mu) {$\bm\mu$};
      \node[const, left=of mu] (Sigma) {$\bm\Sigma$};
      \edge {mu, Sigma} {u};
      \edge {lambda, u} {r};
      \edge {r} {D};
    \end{tikzpicture}
    \label{fig:graphical_model}
  \end{figure}

This gives us our variational parameters $\bm\nu = (\bm\mu, \mathbf{B},
\bm\lambda)^\intercal$. TODO: reference the figure

% TODO: mention in text
%While the ELBO can now be computed without approximations, we ended up imposing
%two zero-variance restrictions, resulting in a less informative model. Moreover,
%we had to abandon the Bayesian treatment of $\bm\lambda$, making the model more
%vulnerable to overfitting.

@@

In this section we derive and simplify the ELBO for this (now fully specified)
model. In order to derive the ELBO, let us go back to Equation \ref{eq:elbo} and
write\footnote{At this point, we will drop the subscript denoting which
  variables the expectation is taken over. Also note that throughout the
  derivation equality is taken to mean `equality up to an additive constant'.}
% TODO: move this to the front and say that the expectations are always w.r.t.
% the full VI approximation unless specified otherwise
\[
  \mathcal{L} = \mathbb{E}[\log\pfull] - \mathbb{E}[\log\approximation].
\]
By substituting in Equations \ref{eq:full} and \ref{eq:approximation}, we get
\begin{align*}
  \mathcal{L} &= \mathbb{E}[\log p(\mathbf{u}) + \log p(\mathbf{r} \mid \mathbf{u}) + \log p(\mathcal{D} \mid \mathbf{r})] \\
              &- \mathbb{E}[\log q(\mathbf{u}) + \log q(\mathbf{r} \mid \mathbf{u})].
\end{align*}
Since $q(\mathbf{r} \mid \mathbf{u}) = p(\mathbf{r} \mid \mathbf{u})$, they
cancel each other out. Also notice that
\begin{align*}
  \mathbb{E}[\log p(\mathbf{u}) - \log q(\mathbf{u})] &= -\DKL{q(\mathbf{u})}{p(\mathbf{u})} \\
                                                      &= -\frac{1}{2} (\tr (\Kuu^{-1}\bm\Sigma) + \bm\mu^\intercal\Kuu^{-1}\bm\mu - m \\
                                                      &+ \log |\Kuu| - \log |\bm\Sigma|),
\end{align*}
by the definition of KL divergence between two multivariate normal distributions
\cite{kl}. Hence,
\begin{align*}
  \mathcal{L} &= \mathbb{E}\left[ \sum_{i=1}^N \sum_{t=1}^T Q_{\mathbf{r}}(s_{i,t}, a_{i,t}) - \V(s_{i,t}) \right] \\
              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
\end{align*}
Using the expressions for $Q_{\mathbf{r}}$ we get
\begin{align*}
  \mathcal{L} &= \mathbb{E}\left[\sum_{i=1}^N \sum_{t=1}^T r(s_{i,t}) - \V(s_{i,t}) + \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\V(s') \right] \\
              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
\end{align*}
We can simplify $\sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t})$ by defining a new vector
$\mathbf{t} = (t_1, \dots, t_{|\mathcal{S}|})^\intercal$, where $t_i$ is the
number of times the state associated with the reward $r_i$ has been visited
across all demonstrations. Then
\begin{align*}
  \mathbb{E} \left[ \sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t}) \right] &= \mathbb{E}[\mathbf{t}^\intercal\mathbf{r}] = \mathbf{t}^\intercal\mathbb{E}[\mathbf{r}] \\
                                                               &= \mathbf{t}^\intercal\mathbb{E}\left[\Kru^\intercal\Kuu^{-1}\mathbf{u}\right] = \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu.
\end{align*}
This allows us to simplify $\mathcal{L}$ to
\begin{align*}
  \mathcal{L} &= \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu - \mathbb{E}[v] \\
              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right),
\end{align*}
where
\[
  v = \sum_{i=1}^N \sum_{t=1}^T \V(s_{i,t}) - \gamma\sum_{s' \in \mathcal{S}}
  \mathcal{T}(s_{i,t}, a_{i,t}, s')\V(s').
\]

@@

The typical way to optimise a quantity (the ELBO, in this case) involves
computing its gradient. Unfortunately, some of the terms in $\mathcal{L}$ are
still left as expected values. The goal of this section is to show how Theorem
\ref{thm:lebesgue} can be applied to our model in order to derive the gradient
anyway\footnote{This technique is inspired by black box variational inference
  \cite{DBLP:conf/aistats/RanganathGB14}, but takes a more detailed look at the
  problem and requires significantly more work to prove correctness.}. After
showing that the theorem applies to our situation, given any term expressed as
an expected value $\mathbb{E}[f(\mathbf{u}, \mathbf{r})]$, we can estimate its
gradient as
\[
  \nabla\mathbb{E}[f(\mathbf{u}, \mathbf{r})] = \mathbb{E}[\nabla f(\mathbf{u},
  \mathbf{r})] \approx \frac{1}{S} \sum_{s=1}^S \nabla f(\mathbf{u}_s,
  \mathbf{r}_s),
\]
which can be computed by drawing $S$ Monte Carlo samples $(\mathbf{u}_s,
\mathbf{r}_s) \sim \approximation$.

Our main goal is Theorem \ref{thm:main}, which allows us to move differentiation
inside the integral. In order to prove it, we use a number of intermediate
results. We start by providing a sketch proof of the measurability of MDP value
functions, which is non-obvious due to their non-trivial definition. We then
prove that the value functions are bounded, and establish four helpful lemmas,
three of which provide well-behaved bounds on derivatives of PDFs. See Figure
\ref{fig:theorems} for an overview of how the results fit together.

@@

\begin{lemma} \label{lemma:integral_of_r}
  \[
    \int \lVert \mathbf{r} \rVert_\infty q(\mathbf{r} \mid \mathbf{u})\,d\mathbf{r} \le a +
    \lVert \Kru^\intercal \Kuu^{-1} \mathbf{u} \rVert_1,
  \]
  where $a$ is a constant independent of $\mathbf{u}$.
\end{lemma}
\begin{proof}
  Since $\rinf \le \lVert \mathbf{r} \rVert_1$,
  \[
    \int \lVert \mathbf{r} \rVert_\infty q(\mathbf{r} \mid \mathbf{u})\,d\mathbf{r} \le \int
    \lVert \mathbf{r} \rVert_1 q(\mathbf{r} \mid \mathbf{u})\,d\mathbf{r} =
    \sum_{i=1}^{|\mathcal{S}|} \mathbb{E}[|r_i|].
  \]
  As each $\mathbb{E}[|r_i|]$ is a mean of a folded Gaussian distribution,
  \[
    \mathbb{E}[|r_i|] = \sigma_i \sqrt{\frac{2}{\pi}} \exp
    \left(-\frac{\xi_i^2}{2\sigma_i^2} \right) + \xi_i \left( 1 - 2\Phi \left(
        -\frac{\xi_1}{\sigma_1} \right) \right),
  \]
  where $\xi_i = \left[\Kru^\intercal\Kuu^{-1}\mathbf{u}\right]_i$, $\sigma_i =
  \sqrt{[\Krr - \Kru^\intercal\Kuu^{-1}\Kru]_{i,i}}$\footnote{The expression
    under the square root sign is non-negative because $\Krr -
    \Kru^\intercal\Kuu^{-1}\Kru$ is a covariance matrix of a Gaussian
    distribution, hence also positive semi-definite, which means that its
    diagonal entries are non-negative.}, and $\Phi$ is the cumulative
  distribution function of the standard normal distribution. Furthermore,
  \[
    \mathbb{E}[|r_i|] \le \sigma_i\sqrt{\frac{2}{\pi}} + |\xi_i|,
  \]
  as $\sigma_i$ is non-negative, and $\Phi(x) \in [0, 1]$ for all $x$. Since
  \[ \sum_{i=1}^{|\mathcal{S}|} |\xi_i| = \lVert \Kru^\intercal \Kuu^{-1}
    \mathbf{u} \rVert_1, \]
  we can set
  \[ a = \sum_{i=1}^{|\mathcal{S}|} \sigma_i \sqrt{\frac{2}{\pi}} \]
  to get the desired result.
\end{proof}

@@

\section{Evaluation} % showing deep understanding

@@

@@

@@

\appendix
\section{Proofs}

>>>>

\subsection{Linear Algebra, Numerical Analysis, and Measure Theory}

Here we introduce a few definitions and results that will be used later in the
paper. Namely, we will use several different vector and matrix norms, consider
how an inverse of a matrix changes with a small perturbation, and use Lebesgue's
dominated convergence theorem in order to justify the validity of our approach.

@@

\section{Solution}
% TODO: express all lambdas as exponentials (update the derivatives or sth)
% TODO: give an overview of this section. Also, consider renaming it or
% splitting it into multiple sections

For any matrix $\mathbf{A}$, we will use either $A_{i,j}$ or
$[\mathbf{A}]_{i,j}$ to denote the element of $\mathbf{A}$ in row $i$ and column
$j$. Moreover, we use $\tr(\mathbf{A})$ to denote its \emph{trace} and
$\adj(\mathbf{A})$ for its \emph{adjugate} (or \emph{classical adjoint}). For
any vector $\mathbf{x}$, we write $\mathbb{R}_d[\mathbf{x}]$ to denote a vector
space of polynomials with degree at most $d$, where variables are elements of
$\mathbf{x}$, and coefficients are in $\mathbb{R}$.

@@

\subsection{Details of the Model}

In order to properly investigate the difference between VI
and ML estimation, we keep other parts of the model the same.
Namely, we set the covariance function to a version of the automatic relevance
detection kernel \cite{DBLP:conf/nips/LevinePK11,neal2012bayesian}
\begin{align*}
  k(\mathbf{x}_i, \mathbf{x}_j) = \lambda_0\exp\left( \vphantom{\frac{1}{2}} \right. &-\frac{1}{2}(\mathbf{x}_i - \mathbf{x}_j)^\intercal\bm\Lambda(\mathbf{x}_i - \mathbf{x}_j) \\
                                                                                     &- \mathbbm{1}[i \ne j]\sigma^2\tr(\bm\Lambda) \left. \vphantom{\frac{1}{2}} \right),
\end{align*}
where $\lambda_0$ is the overall `scale' factor for how similar or distant the
states are, $\bm\Lambda = \diag(\lambda_1, \dots, \lambda_d)$ is a diagonal
matrix that determines the relevance of each feature (where $d$ denotes the
number of features), $\mathbbm{1}$ is defined as
\[
  \mathbbm{1}[b] = \begin{cases}
    1 & \mbox{if $b$ is true} \\
    0 & \mbox{otherwise},
  \end{cases}
\]
and $\sigma^2$ is set to $10^{-2}/2$ (as the original paper noted that
the value makes little difference to the performance of the algorithm
\cite{DBLP:conf/nips/LevinePK11}). We will write $\bm\lambda = (\lambda_0,
\dots, \lambda_d)^\intercal$ to refer to both $\lambda_0$ and $\bm\Lambda$ at
the same time.

% TODO: write about lambda having to be variational
% TODO: mention the open problem of finding the expected value of the inverse of
% a matrix, where each element is a non-trivial function (w.r.t. a parameter of
% the function)
%TODO: introductory sentence for the following assumption:
\begin{equation} \label{eq:approximation}
  \approximation = q(\mathbf{u}) \times q(\mathbf{r} \mid \mathbf{u}),
\end{equation}
where
\begin{equation}
  q(\mathbf{u}) = \mathcal{N}(\mathbf{u}; \bm\mu, \bm\Sigma),
\end{equation}
where $\bm\mu \in \mathbb{R}^{m}$ is the mean vector and the $m \times m$
positive semi-definite matrix $\bm\Sigma$ is the covariance matrix
\cite{DBLP:conf/nips/ChengB17,DBLP:journals/jmlr/HensmanDS17,DBLP:conf/aaai/HoangHL17}.

Ong et al. \cite{ong2018gaussian} have recently suggested that, in order to make
variational approximation of a multivariate Gaussian more scalable, the
covariance matrix should be decomposed as $\bm\Sigma =
\mathbf{B}\mathbf{B}^\intercal + \mathbf{D}^2$, where $\mathbf{B}$ is a lower
triangular $m \times p$ matrix with positive diagonal entries, and $\mathbf{D}$
is a diagonal matrix. Typically, we would set $p$ so that $p \ll m$ to get an
efficient approximation. However, as our goal is precision rather than
scalability, we will set $p = m$ and $\mathbf{D} = \mathbf{O}_m$ in order to
retain full covariance structure.

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node[latent] (u) {$\mathbf{u}$};
    \node[latent, right=of u] (r) {$\mathbf{r}$};
    \node[obs, right=of r] (D) {$\mathcal{D}$};
    \node[const, below=of r] (lambda) {$\bm\lambda$};
    \node[const, below=of u] (mu) {$\bm\mu$};
    \node[const, left=of mu] (B) {$\mathbf{B}$};
    \edge {mu, B} {u};
    \edge {lambda, u} {r};
    \edge {r} {D};
  \end{tikzpicture}
  \caption{Our VI problem expressed as a (simplified)
    Bayesian network. The only observed variable (representing the
    demonstrations) is in a gray circle, modelled latent variables are in white
    circles, and the variational parameters are at the bottom.}
  \label{fig:graphical_model}
\end{figure}

The resulting model is summarised in Figure \ref{fig:graphical_model}. We rely
on $p(\mathcal{D} \mid \mathbf{r})$ as the only link between data and the model.
Since the expression for $q(\mathbf{r} \mid \mathbf{u})$ has both $\mathbf{u}$
and covariance matrices in it, $\mathbf{r}$ depends on both $\mathbf{u}$ and the
parameters of the kernel, $\bm\lambda$. The two remaining dependencies are quite
clear, given that the approximating distribution for $\mathbf{u}$ is
$\mathcal{N}(\bm\lambda, \mathbf{BB}^\intercal)$. % TODO: concluding sentence

@@

In this section, we derive and simplify the ELBO for this (now fully specified)
model. Note that in order to keep the derivation simple, we drop all constant
terms in the expression of $\mathcal{L}$, i.e., equality is taken to mean
`equality up to an additive constant'. Also note that all expected values are
with respect to $(\mathbf{u}, \mathbf{r}) \sim q(\mathbf{u}, \mathbf{r})$.

In order to derive the ELBO, let us go back to Equation \ref{eq:elbo} and
write
\[
  \mathcal{L} = \mathbb{E}[\log\pfull] - \mathbb{E}[\log\approximation].
\]
By substituting in Equations \ref{eq:full} and \ref{eq:approximation}, we get
\begin{align*}
  \mathcal{L} &= \mathbb{E}[\log p(\mathbf{u}) + \log p(\mathbf{r} \mid \mathbf{u}) + \log p(\mathcal{D} \mid \mathbf{r})] \\
              &- \mathbb{E}[\log q(\mathbf{u}) + \log q(\mathbf{r} \mid \mathbf{u})].
\end{align*}
Since $q(\mathbf{r} \mid \mathbf{u}) = p(\mathbf{r} \mid \mathbf{u})$, they
cancel each other out. Also notice that
\begin{align*}
  \mathbb{E}[\log p(\mathbf{u}) - \log q(\mathbf{u})] &= -\DKL{q(\mathbf{u})}{p(\mathbf{u})} \\
                                                      &= -\frac{1}{2} (\tr (\Kuu^{-1}\bm\Sigma) + \bm\mu^\intercal\Kuu^{-1}\bm\mu - m \\
                                                      &+ \log |\Kuu| - \log |\bm\Sigma|),
\end{align*}
by the definition of KL divergence between two multivariate Gaussians
\cite{kl}. Hence,
\begin{align*}
  \mathcal{L} &= \mathbb{E}\left[ \sum_{i=1}^N \sum_{t=1}^T Q_{\mathbf{r}}(s_{i,t}, a_{i,t}) - \V(s_{i,t}) \right] \\
              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
\end{align*}
Using the expressions for $Q_{\mathbf{r}}$ we get
\begin{align*}
  \mathcal{L} &= \mathbb{E}\left[\sum_{i=1}^N \sum_{t=1}^T r(s_{i,t}) - \V(s_{i,t}) + \gamma\sum_{s' \in \mathcal{S}} \mathcal{T}(s_{i,t}, a_{i,t}, s')\V(s') \right] \\
              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right).
\end{align*}
We can simplify $\sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t})$ by defining a new vector
$\mathbf{t} = (t_1, \dots, t_{|\mathcal{S}|})^\intercal$, where $t_i$ is the
number of times the state associated with the reward $r_i$ has been visited
across all demonstrations. Then
\begin{align*}
  \mathbb{E} \left[ \sum_{i=1}^N\sum_{t=1}^Tr(s_{i,t}) \right] &= \mathbb{E}[\mathbf{t}^\intercal\mathbf{r}] = \mathbf{t}^\intercal\mathbb{E}[\mathbf{r}] \\
                                                               &= \mathbf{t}^\intercal\mathbb{E}\left[\Kru^\intercal\Kuu^{-1}\mathbf{u}\right] = \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu.
\end{align*}
This allows us to simplify $\mathcal{L}$ to
\begin{align*}
  \mathcal{L} &= \mathbf{t}^\intercal\Kru^\intercal\Kuu^{-1}\bm\mu - \mathbb{E}[v] \\
              &- \frac{1}{2} \left(\tr \left( \Kuu^{-1}\bm\Sigma \right) + \bm\mu^\intercal\Kuu^{-1}\bm\mu + \log |\Kuu| - \log |\bm\Sigma| \right),
\end{align*}
where
\[
  v = \sum_{i=1}^N \sum_{t=1}^T \V(s_{i,t}) - \gamma\sum_{s' \in \mathcal{S}}
  \mathcal{T}(s_{i,t}, a_{i,t}, s')\V(s').
\]

@@

The typical way to optimise a quantity (the ELBO, in this case) involves
computing its gradient. Unfortunately, some of the terms in $\mathcal{L}$ are
still left as expected values. The goal of this section is to show how Theorem
\ref{thm:lebesgue} can be applied to our model in order to derive the gradient
anyway\footnote{This technique is inspired by black box VI
  \cite{DBLP:conf/aistats/RanganathGB14}, but takes a more detailed look at the
  problem and requires significantly more work to prove correctness.}. After
showing that the theorem applies to our situation, given any term expressed as
an expected value $\mathbb{E}[f(\mathbf{u}, \mathbf{r})]$, we can estimate its
gradient as
\[
  \nabla\mathbb{E}[f(\mathbf{u}, \mathbf{r})] = \mathbb{E}[\nabla f(\mathbf{u},
  \mathbf{r})] \approx \frac{1}{S} \sum_{s=1}^S \nabla f(\mathbf{u}_s,
  \mathbf{r}_s),
\]
which can be computed by drawing $S$ Monte Carlo samples $(\mathbf{u}_s,
\mathbf{r}_s) \sim \approximation$.

Our main goal is Theorem \ref{thm:main}, which allows us to move differentiation
inside the integral. In order to prove it, we use a number of intermediate
results. We start by stating a few derivatives of probability density functions
(PDFs) and covariance matrices, and bound their values with some
easy-to-deal-with polynomials. We then provide a sketch proof of the
measurability of MDP value functions, which is non-obvious due to their
non-trivial definition. Afterwards, we establish bounds for the value functions,
and, after another quick lemma, tackle the main proof of this paper. See Figure
\ref{fig:theorems} for an overview of how these results fit together.

Before that, however, we define a few extra variables in order to simplify
expressions of derivatives:
\begin{align*}
  \mathbf{U} &= (\mathbf{u} - \bm\mu)(\mathbf{u} - \bm\mu)^\intercal, \\
  \mathbf{S} &= \Kru^\intercal\Kuu^{-1}, \\
  \bm\Gamma &= \Krr - \mathbf{S}\Kru, \\
  \mathbf{R} &= \mathbf{S}\frac{\partial \Kru}{\partial \lambda_i} - \frac{\partial \Krr}{\partial \lambda_i} + \left( \frac{\partial \Kru^\intercal}{\partial \lambda_i} - \mathbf{S}\frac{\partial \Kuu}{\partial \lambda_i} \right) \Kuu^{-1}\Kru, \\
  Q &= (\mathbf{u} - \bm\mu)^\intercal\bm\Sigma^{-1}(\mathbf{u} - \bm\mu).
\end{align*}

\begin{restatable}[Derivatives of PDFs]{lemma}{derivatives} \label{lemma:derivatives}
  \leavevmode
  \begin{enumerate}
  \item $\frac{\partial q(\mathbf{u})}{\partial \bm\mu} =
    \frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-1} + \bm\Sigma^{-\intercal})(\mathbf{u}
    - \bm\mu)$.
  \item
    \begin{enumerate}
    \item
      $\frac{\partial q(\mathbf{u})}{\partial \bm\Sigma} =
      \frac{1}{2}q(\mathbf{u})(\bm\Sigma^{-\intercal}\mathbf{U}\bm\Sigma^{-\intercal}
      - \bm\Sigma^{-\intercal})$.
    \item
      $\frac{\partial q(\mathbf{u})}{\partial \mathbf{B}} =
      q(\mathbf{u})(\bm\Sigma^{-1}\mathbf{U}\bm\Sigma^{-1} -
      |\bm\Sigma|^{-1}\adj(\bm\Sigma))\mathbf{B}$.
    \end{enumerate}
  \item For $i = 0, \dots, d$,
    \begin{enumerate}
    \item
      \begin{align*}
        \frac{\partial q(\mathbf{r} \mid \mathbf{u})}{\partial \lambda_i} &= \frac{1}{2}q(\mathbf{r} \mid \mathbf{u}) (|\bm\Gamma|^{-1} \tr(\mathbf{R} \adj(\bm\Gamma)) \\
                                                                          &- (\mathbf{r} - \mathbf{Su})^\intercal\bm\Gamma^{-1}\mathbf{R}\bm\Gamma^{-1}(\mathbf{r} - \mathbf{Su})).
      \end{align*}
    \item For any covariance matrix $\mathbf{K}$,
      \[
        \frac{\partial \mathbf{K}}{\partial \lambda_i} =
        \begin{cases}
          \frac{1}{\lambda_i}\mathbf{K} & \text{if } i = 0, \\
          \mathbf{L} & \text{otherwise,}
        \end{cases}
      \]
      where
      \[
        L_{j,k} = k(\mathbf{x}_j, \mathbf{x}_k) \left( -\frac{1}{2}(x_{j,i} -
          x_{k,i})^2 - \mathbbm{1}[j \ne k]\sigma^2 \right).
      \]
    \end{enumerate}
  \end{enumerate}
\end{restatable}

@@

\begin{lemma} \label{lemma:integral_of_r}
  \[
    \int \lVert \mathbf{r} \rVert_\infty q(\mathbf{r} \mid \mathbf{u})\,d\mathbf{r} \le a +
    \lVert \Kru^\intercal \Kuu^{-1} \mathbf{u} \rVert_1,
  \]
  where $a$ is a constant independent of $\mathbf{u}$.
\end{lemma}
\begin{proof}
  Since $\rinf \le \lVert \mathbf{r} \rVert_1$,
  \[
    \int \lVert \mathbf{r} \rVert_\infty q(\mathbf{r} \mid \mathbf{u})\,d\mathbf{r} \le \int
    \lVert \mathbf{r} \rVert_1 q(\mathbf{r} \mid \mathbf{u})\,d\mathbf{r} =
    \sum_{i=1}^{|\mathcal{S}|} \mathbb{E}[|r_i|].
  \]
  As each $\mathbb{E}[|r_i|]$ is a mean of a folded Gaussian distribution,
  \[
    \mathbb{E}[|r_i|] = \sigma_i \sqrt{\frac{2}{\pi}} \exp
    \left(-\frac{\xi_i^2}{2\sigma_i^2} \right) + \xi_i \left( 1 - 2\Phi \left(
        -\frac{\xi_1}{\sigma_1} \right) \right),
  \]
  where $\xi_i = \left[\Kru^\intercal\Kuu^{-1}\mathbf{u}\right]_i$, $\sigma_i =
  \sqrt{[\Krr - \Kru^\intercal\Kuu^{-1}\Kru]_{i,i}}$\footnote{The expression
    under the square root sign is non-negative because $\Krr -
    \Kru^\intercal\Kuu^{-1}\Kru$ is a covariance matrix of a Gaussian
    distribution, hence also positive semi-definite, which means that its
    diagonal entries are non-negative.}, and $\Phi$ is the cumulative
  distribution function of the standard Gaussian. Furthermore,
  \[
    \mathbb{E}[|r_i|] \le \sigma_i\sqrt{\frac{2}{\pi}} + |\xi_i|,
  \]
  as $\sigma_i$ is non-negative, and $\Phi(x) \in [0, 1]$ for all $x$. Since
  \[ \sum_{i=1}^{|\mathcal{S}|} |\xi_i| = \lVert \Kru^\intercal \Kuu^{-1}
    \mathbf{u} \rVert_1, \]
  we can set
  \[ a = \sum_{i=1}^{|\mathcal{S}|} \sigma_i \sqrt{\frac{2}{\pi}} \]
  to get the desired result.
\end{proof}

@@

\section{Evaluation} % showing deep understanding
% Could measure how algorithms scale with the size of the map (number of states)
% Could add plots that show how the ELBO changes with different parameter values
% (the inverse parabolas)
% And plot you get if you start with mu = 0 and move to negative infinity (hint:
% it increases)

@@

More variables/information. Fewer assumptions.

We show how to avoid the deterministic training conditional assumption.

@@

\begin{itemize}
\item Let the agent take uncertainty (variation estimates) into account when
  choosing what to do next (i.e., choosing what to learn).
\item MDP solver that works with distributions of rewards.
\item Stress how solving MDPs is the performance bottleneck: num_samples MDPs
  have to be solved for each iteration.
  \begin{itemize}
  \item Integration with MDP solvers that approximate the value function
    sufficiently well for training.
  \item Or perhaps there is a way to ``modify'' a solution to one MDP in order
    to get a solution for a similar MDP (there was a FATA talk about that, but
    for a different problem of course).
  \item Perhaps value iteration converges faster if the values are initialised
    to the optimal values of the previous iteration.
  \end{itemize}
\end{itemize}

@@

\appendix
\section{Proofs} \label{appendix:proofs}
